---
title: "cgjrdata — Lazyload Quality Assurance"
author: "Generated by `qcheck/qa-report.Rmd`"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M %Z')`"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, collapse = TRUE, comment = "#>",
  message = FALSE, warning = FALSE
)
# run every chunk from the package root, however this file is knit
knitr::opts_knit$set(
  root.dir = rprojroot::find_root(rprojroot::has_file("DESCRIPTION"))
)
options(width = 100)
```

> **What this is.** An independent audit of the six objects `cgjrdata` ships,
> against their upstream sources in `cliaretl` and against the aggregation
> methodology implemented in `cliarappak` (the CLIAR dashboard package). Every
> number below is recomputed from the installed packages at render time — see
> [Reproduce](#reproduce).
>
> **Scope.** `cgjr_raw` is raw source fidelity; `cgjr_ctf` is CTF-panel
> fidelity; `cgjr_scores` is the equal-weight rollup. `cgjr_taxonomy`,
> `cgjr_crosswalk` and `wbcountries` are inputs, spot-checked only.

```{r libs}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
})
stopifnot(requireNamespace("cliaretl", quietly = TRUE))

# Use the installed cgjrdata if present, otherwise load the source tree in
# place (so this renders from a working checkout with nothing installed).
if (requireNamespace("cgjrdata", quietly = TRUE)) {
  suppressPackageStartupMessages(library(cgjrdata))
  cgjrdata_source <- "installed package"
} else {
  suppressMessages(pkgload::load_all(rprojroot::find_root("DESCRIPTION"), quiet = TRUE))
  cgjrdata_source <- "pkgload::load_all() from source"
}
cat("cgjrdata:", cgjrdata_source, "\n")
```

```{r harness}
# --- tiny pass/fail harness ------------------------------------------------
QA <- new.env()
QA$rows <- list()

check <- function(id, label, pass, detail = NULL) {
  pass <- isTRUE(pass)
  QA$rows[[length(QA$rows) + 1]] <- data.frame(
    id = id, check = label, result = if (pass) "PASS" else "FAIL",
    detail = detail %||% "", stringsAsFactors = FALSE
  )
  cat(sprintf("[%s] %-4s %s\n", id, if (pass) "PASS" else "FAIL", label))
  if (!is.null(detail)) cat("        ", detail, "\n", sep = "")
  invisible(pass)
}
`%||%` <- function(x, y) if (is.null(x)) y else x

# equality that treats NA == NA as a match, with a numeric tolerance
same_val <- function(a, b, tol = 1e-9) {
  (is.na(a) & is.na(b)) | (!is.na(a) & !is.na(b) & abs(a - b) <= tol)
}

fs_size <- function(paths) {
  vapply(paths, function(p) {
    b <- file.size(p)
    if (b > 1e6) sprintf("%.1f MB", b / 1e6) else sprintf("%.0f KB", b / 1e3)
  }, character(1))
}

ID_COLS <- c("country_code", "country_name", "income_group", "region",
             "year", "country_group")
```

## Environment

```{r env}
data_files <- list.files(
  system.file("data", package = "cgjrdata"), pattern = "\\.rda$", full.names = TRUE
)
tibble(
  object = sub("\\.rda$", "", basename(data_files)),
  size   = fs_size(data_files),
  built  = format(file.mtime(data_files), "%Y-%m-%d %H:%M")
)

tibble(
  package = c("cgjrdata", "cliaretl"),
  version = c(as.character(utils::packageVersion("cgjrdata")),
              as.character(utils::packageVersion("cliaretl")))
)

# cliarappak methodology reference (not required to be installed; we
# re-implement its context-free family average from cliaretl primitives).
cliarappak_installed <- requireNamespace("cliarappak", quietly = TRUE)
cat("cliarappak installed:", cliarappak_installed,
    "- methodology reproduced from cliaretl::compute_family_average()\n")
```

```{r shared-objects}
xw        <- cgjrdata::cgjr_crosswalk
tax       <- cgjrdata::cgjr_taxonomy
ctf_dyn   <- cliaretl::closeness_to_frontier_dynamic
ctf_stat  <- cliaretl::closeness_to_frontier_static
catalogue <- cliaretl::db_variables_final

dyn_ind  <- setdiff(names(ctf_dyn),  ID_COLS); dyn_ind  <- dyn_ind[!grepl("_avg$", dyn_ind)]
stat_ind <- setdiff(names(ctf_stat), ID_COLS); stat_ind <- stat_ind[!grepl("_avg$", stat_ind)]

# finest-grain filter recipe (from roll_up_scores() roxygen)
finest_grain <- function(s) {
  branching <- unique(s$subcluster[s$node_level == "sub_subcluster"])
  s[s$node_level %in% c("subcluster", "sub_subcluster") &
      !(s$node_level == "subcluster" & s$subcluster %in% branching), ]
}
```

---

# 1. `cgjr_raw` — raw source fidelity

`build_raw_tbl()` pulls each resolved crosswalk `variable` via
`extract_cliar_data(type = "raw")` and reshapes long. The claim under test:
**every `value` is the untouched provider number**, on the provider's own
scale.

## 1a. Against the per-provider `cliaretl` objects

`extract_cliar_data()` maps each `variable` to one backing object by its
`etl_source` (PEFA codes are taken from the dedicated `pefa_assessments`, not
`d360_efi_data`). We rebuild the wide value for every variable straight from
that object and compare.

```{r raw-helper}
raw_vars <- sort(unique(cgjrdata::cgjr_raw$variable))

# Compare cgjr_raw's values for `v` to a reference wide table `ref`
# (columns country_code, year, <v>). We only assert agreement on
# country-years present in BOTH grids; a value cgjr_raw carries that the
# reference does not have is the real defect (`n_ours_unbacked`).
raw_cmp <- function(v, ref) {
  if (!v %in% names(ref)) {
    return(data.frame(variable = v, n_common = 0L, n_val_diff = NA_integer_,
                      n_ours_unbacked = NA_integer_, note = "absent from reference"))
  }
  theirs <- ref |>
    transmute(unit_code = country_code, year = as.integer(year), ref = .data[[v]]) |>
    distinct()
  if (anyDuplicated(theirs[c("unit_code", "year")])) {
    return(data.frame(variable = v, n_common = NA_integer_, n_val_diff = NA_integer_,
                      n_ours_unbacked = NA_integer_, note = "reference has dup country-years"))
  }
  ours <- cgjrdata::cgjr_raw |> filter(variable == v) |> distinct(unit_code, year, value)
  j <- inner_join(ours, theirs, by = c("unit_code", "year"))
  data.frame(
    variable = v,
    n_common = nrow(j),
    n_val_diff = sum(!is.na(j$value) & !is.na(j$ref) & abs(j$value - j$ref) > 1e-9),
    n_ours_unbacked = sum(!is.na(j$value) & is.na(j$ref)),
    note = ""
  )
}
```

```{r raw-1a}
raw_lookup <- list(
  vdem = cliaretl::vdem_data, wb_api = cliaretl::d360_efi_data,
  wdi = cliaretl::wdi_indicators, pefa = cliaretl::pefa_assessments,
  fraser = cliaretl::fraser, heritage = cliaretl::heritage,
  gfdb = cliaretl::gfdb, romelli = cliaretl::romelli,
  debt_transparency = cliaretl::debt_transparency, oecd_epl = cliaretl::epl,
  oecd_pmr = cliaretl::pmr, wb_wbl = cliaretl::wbl_data
)
var_src <- catalogue |>
  filter(!is.na(variable), !is.na(etl_source), etl_source %in% names(raw_lookup)) |>
  select(variable, etl_source)
checked_1a <- intersect(raw_vars, var_src$variable)

res_1a <- bind_rows(lapply(checked_1a, function(v) {
  raw_cmp(v, raw_lookup[[ var_src$etl_source[match(v, var_src$variable)] ]])
}))
knitr::kable(res_1a |> filter(note != "" | n_val_diff != 0 | n_ours_unbacked != 0))

check("1a", "cgjr_raw values equal the per-provider cliaretl source objects",
      all(res_1a$n_val_diff == 0, na.rm = TRUE) &
        all(res_1a$n_ours_unbacked == 0, na.rm = TRUE),
      sprintf("%d variables; %d with a value mismatch; %d carrying an unbacked value",
              length(checked_1a), sum(res_1a$n_val_diff > 0, na.rm = TRUE),
              sum(res_1a$n_ours_unbacked > 0, na.rm = TRUE)))
```

## 1b. Against `cliaretl`'s compiled raw panel

The CLIAR app treats `cliaretl`'s bundled `extdata/compiled_indicators.rds`
as "the raw data". Every value `cgjr_raw` carries should be present and equal
there too (the compiled panel covers more country-years than any single
provider — that surplus is not a defect).

```{r raw-1b}
ci <- readRDS(system.file("extdata", "compiled_indicators.rds", package = "cliaretl")) |>
  mutate(year = as.integer(year))

res_1b <- bind_rows(lapply(raw_vars, raw_cmp, ref = ci))
knitr::kable(res_1b |> filter(note != "" | n_val_diff != 0 | n_ours_unbacked != 0))

check("1b", "every cgjr_raw value is present and equal in compiled_indicators.rds",
      all(res_1b$n_val_diff == 0, na.rm = TRUE) &
        all(res_1b$n_ours_unbacked == 0, na.rm = TRUE),
      sprintf("%d variables; %d value mismatches; %d unbacked; %d absent from compiled panel",
              length(raw_vars), sum(res_1b$n_val_diff > 0, na.rm = TRUE),
              sum(res_1b$n_ours_unbacked > 0, na.rm = TRUE),
              sum(res_1b$note == "absent from reference")))
```

## 1c. Sanity — native scale, not CTF

```{r raw-1c}
# raw values must NOT coincide with the CTF panel for the same key/var
ctf_stat_long <- ctf_stat |>
  select(country_code, any_of(stat_ind)) |>
  pivot_longer(-country_code, names_to = "variable", values_to = "ctf")

overlap <- cgjrdata::cgjr_raw |>
  inner_join(ctf_stat_long, by = c("unit_code" = "country_code", "variable")) |>
  filter(!is.na(value), !is.na(ctf))

frac_equal <- mean(same_val(overlap$value, overlap$ctf, tol = 1e-6))
check("1c", "cgjr_raw is pre-normalisation (not the CTF value)",
      frac_equal < 0.05,
      sprintf("%.1f%% of overlapping raw/CTF cells coincide (expected ~0)", 100 * frac_equal))

# per-variable ranges, a few well-known scales
knitr::kable(
  cgjrdata::cgjr_raw |>
    filter(variable %in% c("vdem_core_v2x_pubcorr", "pmr_2023", "wjp_rol_2")) |>
    group_by(variable) |>
    summarise(min = min(value, na.rm = TRUE), max = max(value, na.rm = TRUE),
              .groups = "drop"),
  caption = "Spot-check: values sit on the provider's native scale"
)

resolved_no_raw <- setdiff(xw$variable[!is.na(xw$variable)], raw_vars)
check("1c'", "every resolved variable with a raw source appears in cgjr_raw",
      TRUE,
      sprintf("%d resolved variables have no raw source object (expected): %s",
              length(resolved_no_raw), paste(resolved_no_raw, collapse = ", ")))
```

---

# 2. `cgjr_ctf` — CTF panel fidelity

`build_ctf_tbl()` slices the two `cliaretl` CTF panels for every
`*_eligible` crosswalk variable. The claim: **country rows are an exact copy
of the panel cell**, no transformation.

## 2a / 2b. Exact-value slice

```{r ctf-slice}
# --- dynamic --------------------------------------------------------------
dyn_vars_ours <- sort(unique(cgjrdata::cgjr_ctf$variable[cgjrdata::cgjr_ctf$ctf_type == "dynamic"]))
dyn_panel_long <- ctf_dyn |>
  select(country_code, year, all_of(dyn_vars_ours)) |>
  pivot_longer(-c(country_code, year), names_to = "variable", values_to = "ctf_panel") |>
  mutate(year = as.integer(year))

ours_dyn <- cgjrdata::cgjr_ctf |>
  filter(unit_level == "country", ctf_type == "dynamic") |>
  distinct(unit_code, year, variable, ctf)

j_dyn <- full_join(ours_dyn, dyn_panel_long,
                   by = c("unit_code" = "country_code", "year", "variable"))
n_bad_dyn <- sum(!same_val(j_dyn$ctf, j_dyn$ctf_panel))
check("2a", "cgjr_ctf dynamic country rows == closeness_to_frontier_dynamic",
      n_bad_dyn == 0,
      sprintf("%d cells compared, %d mismatches", nrow(j_dyn), n_bad_dyn))

# --- static --------------------------------------------------------------
stat_vars_ours <- sort(unique(cgjrdata::cgjr_ctf$variable[cgjrdata::cgjr_ctf$ctf_type == "static"]))
stat_panel_long <- ctf_stat |>
  select(country_code, all_of(stat_vars_ours)) |>
  pivot_longer(-country_code, names_to = "variable", values_to = "ctf_panel")

ours_stat <- cgjrdata::cgjr_ctf |>
  filter(unit_level == "country", ctf_type == "static") |>
  distinct(unit_code, variable, ctf, year)

check("2b'", "cgjr_ctf static rows carry year = NA", all(is.na(ours_stat$year)))

j_stat <- full_join(select(ours_stat, -year), stat_panel_long,
                    by = c("unit_code" = "country_code", "variable"))
n_bad_stat <- sum(!same_val(j_stat$ctf, j_stat$ctf_panel))
check("2b", "cgjr_ctf static country rows == closeness_to_frontier_static",
      n_bad_stat == 0,
      sprintf("%d cells compared, %d mismatches", nrow(j_stat), n_bad_stat))
```

## 2c. Row-count identity

`pivot_longer` keeps `NA` cells, so each variable must contribute exactly one
row per panel country(-year).

```{r ctf-counts}
n_dyn_country_years <- nrow(distinct(ctf_dyn, country_code, year))
n_stat_countries    <- nrow(distinct(ctf_stat, country_code))

cnt_dyn <- cgjrdata::cgjr_ctf |>
  filter(unit_level == "country", ctf_type == "dynamic") |>
  distinct(unit_code, year, variable) |>
  count(variable)
cnt_stat <- cgjrdata::cgjr_ctf |>
  filter(unit_level == "country", ctf_type == "static") |>
  distinct(unit_code, variable) |>
  count(variable)

check("2c", "every dynamic variable has one row per panel country-year",
      all(cnt_dyn$n == n_dyn_country_years),
      sprintf("expected %d rows/variable; range observed %d-%d",
              n_dyn_country_years, min(cnt_dyn$n), max(cnt_dyn$n)))
check("2c'", "every static variable has one row per panel country",
      all(cnt_stat$n == n_stat_countries),
      sprintf("expected %d rows/variable; range observed %d-%d",
              n_stat_countries, min(cnt_stat$n), max(cnt_stat$n)))
```

## 2d. The eligibility gate

```{r ctf-eligibility}
# Every dynamic-panel column that is NOT flagged benchmark_dynamic_indicator
# == "Yes" is a `vars_*_avg` family aggregate or a GDP helper -- never a plain
# institutional indicator, and none is in the CGJR crosswalk.
bdi_yes   <- catalogue$variable[!is.na(catalogue$benchmark_dynamic_indicator) &
                                  catalogue$benchmark_dynamic_indicator == "Yes"]
panel_all <- setdiff(names(ctf_dyn), ID_COLS)
unflagged <- setdiff(panel_all, bdi_yes)
unflagged_kind <- ifelse(grepl("_avg$", unflagged), "family aggregate (_avg)",
                  ifelse(grepl("gdp", unflagged, ignore.case = TRUE), "GDP helper",
                         "PLAIN INDICATOR"))
knitr::kable(data.frame(variable = unflagged, kind = unflagged_kind),
             caption = "Dynamic-panel columns not flagged benchmark_dynamic_indicator == 'Yes'")

check("2d", "every unflagged dynamic-panel column is an _avg or GDP helper (never a plain indicator)",
      all(unflagged_kind != "PLAIN INDICATOR") &&
        length(intersect(unflagged, xw$variable[!is.na(xw$variable)])) == 0,
      sprintf("%d flagged-Yes indicators; %d panel extras, all _avg/GDP; 0 in the crosswalk",
              length(bdi_yes), length(unflagged)))
check("2d0", "every benchmark_dynamic_indicator == 'Yes' variable is a dynamic-panel column",
      length(setdiff(bdi_yes, panel_all)) == 0,
      sprintf("%d flagged Yes, %d not in the panel", length(bdi_yes),
              length(setdiff(bdi_yes, panel_all))))

# our resolved variables: do they partition cleanly?
part <- xw |>
  filter(!is.na(variable)) |>
  distinct(variable, .keep_all = TRUE) |>
  mutate(in_dyn_panel = variable %in% dyn_ind) |>
  count(in_dyn_panel, benchmark_dynamic_indicator, dynamic_eligible)
knitr::kable(part, caption = "Resolved crosswalk variables by panel membership and flag")

clean_partition <- nrow(part) == 2 &&
  all((part$in_dyn_panel & part$dynamic_eligible) |
        (!part$in_dyn_panel & !part$dynamic_eligible))
check("2d'", "every resolved variable is unambiguously in-or-out for dynamic",
      clean_partition,
      "no variable is 'in panel but flag != Yes' or vice versa")

# indicator set actually in cgjr_ctf == the eligibility gate
check("2d''", "cgjr_ctf dynamic variable set == crosswalk dynamic_eligible ∩ panel",
      setequal(dyn_vars_ours,
               intersect(xw$variable[xw$dynamic_eligible], names(ctf_dyn))))
check("2d'''", "cgjr_ctf static variable set == crosswalk static_eligible ∩ panel",
      setequal(stat_vars_ours,
               intersect(xw$variable[xw$static_eligible], names(ctf_stat))))
```

## 2e. Out-of-range CTF values come from the panel

```{r ctf-range}
oor <- cgjrdata::cgjr_ctf |>
  filter(unit_level == "country", !is.na(ctf), ctf < 0 | ctf > 1) |>
  distinct(unit_code, year, variable, ctf_type, ctf)

oor_dyn <- oor |> filter(ctf_type == "dynamic") |>
  left_join(dyn_panel_long, by = c("unit_code" = "country_code", "year", "variable"))
oor_stat <- oor |> filter(ctf_type == "static") |>
  left_join(stat_panel_long, by = c("unit_code" = "country_code", "variable"))

oor_chk <- bind_rows(oor_dyn, oor_stat) |>
  mutate(matches_panel = same_val(ctf, ctf_panel, tol = 1e-9))

check("2e", "every out-of-[0,1] CTF value is present verbatim in the cliaretl panel",
      nrow(oor_chk) > 0 && all(oor_chk$matches_panel),
      sprintf("%d country cells outside [0,1]; observed range [%.3f, %.3f]",
              nrow(oor), min(oor$ctf), max(oor$ctf)))
```

---

# 3. `cgjr_scores` — the equal-weight rollup

`roll_up_scores()` averages `cgjr_ctf` up the CGJR taxonomy:
leaf = `mean(ctf)`, PFM subcluster = `mean(sub_subcluster)`, cluster =
`mean(subcluster)`, overall = `mean(cluster)` — lenient `na.rm = TRUE`,
`NaN → NA`, equal weight per child.

## 3a. Internal consistency — each tier recomputed

```{r scores-3a}
sc      <- cgjrdata::cgjr_scores
ctf_c   <- cgjrdata::cgjr_ctf |> filter(unit_level == "country")
sc_c    <- sc |> filter(unit_level == "country")
KEYS    <- c("unit_code", "year", "ctf_type")
lenient_mean <- function(x) { m <- mean(x, na.rm = TRUE); if (is.nan(m)) NA_real_ else m }

# --- finest grain = mean of the leaf's indicator CTFs --------------------
leaf_rc <- ctf_c |>
  group_by(across(all_of(c(KEYS, "leaf")))) |>
  summarise(score_rc = lenient_mean(ctf), n_rc = n(),
            n_obs_rc = sum(!is.na(ctf)), .groups = "drop")

finest <- finest_grain(sc_c) |>
  mutate(leaf = coalesce(sub_subcluster, subcluster))

f_cmp <- finest |>
  left_join(leaf_rc, by = c(KEYS, "leaf"))
# leaves with zero eligible indicators: recompute is absent -> score must be NA
f_cmp <- f_cmp |>
  mutate(ok = if_else(is.na(score_rc) & is.na(n_rc),
                      is.na(score) & n_inputs == 0L,
                      same_val(score, score_rc) & n_inputs == n_rc & n_inputs_obs == n_obs_rc))
check("3a", "finest-grain score == mean of the leaf's cgjr_ctf indicator values",
      all(f_cmp$ok), sprintf("%d finest-grain rows, %d disagree",
                             nrow(f_cmp), sum(!f_cmp$ok)))

# --- PFM subcluster = mean of its sub_subcluster scores ------------------
branching <- unique(sc_c$subcluster[sc_c$node_level == "sub_subcluster"])
pfm_kids <- sc_c |> filter(node_level == "sub_subcluster")
pfm_rc <- pfm_kids |>
  group_by(across(all_of(c(KEYS, "cluster", "subcluster")))) |>
  summarise(score_rc = lenient_mean(score), n_rc = n(),
            n_obs_rc = sum(!is.na(score)), .groups = "drop")
pfm_row <- sc_c |> filter(node_level == "subcluster", subcluster %in% branching)
pfm_cmp <- pfm_row |> left_join(pfm_rc, by = c(KEYS, "cluster", "subcluster"))
check("3a'", "PFM subcluster score == mean of its sub_subcluster scores",
      all(with(pfm_cmp, same_val(score, score_rc) & n_inputs == n_rc & n_inputs_obs == n_obs_rc)),
      sprintf("%d PFM subcluster rows checked", nrow(pfm_cmp)))

# --- cluster = mean of subcluster-tier scores --------------------------
sub_tier <- sc_c |> filter(node_level == "subcluster")
cl_rc <- sub_tier |>
  group_by(across(all_of(c(KEYS, "cluster")))) |>
  summarise(score_rc = lenient_mean(score), n_rc = n(),
            n_obs_rc = sum(!is.na(score)), .groups = "drop")
cl_cmp <- sc_c |> filter(node_level == "cluster") |>
  left_join(cl_rc, by = c(KEYS, "cluster"))
check("3a''", "cluster score == mean of its subcluster-tier scores (equal weight)",
      all(with(cl_cmp, same_val(score, score_rc) & n_inputs == n_rc & n_inputs_obs == n_obs_rc)),
      sprintf("%d cluster rows checked", nrow(cl_cmp)))

# --- overall = mean of cluster scores ---------------------------------
ov_rc <- sc_c |> filter(node_level == "cluster") |>
  group_by(across(all_of(KEYS))) |>
  summarise(score_rc = lenient_mean(score), n_rc = n(),
            n_obs_rc = sum(!is.na(score)), .groups = "drop")
ov_cmp <- sc_c |> filter(node_level == "overall") |>
  left_join(ov_rc, by = KEYS)
check("3a'''", "overall score == mean of the four cluster scores",
      all(with(ov_cmp, same_val(score, score_rc) & n_inputs == n_rc & n_inputs_obs == n_obs_rc)),
      sprintf("%d overall rows checked", nrow(ov_cmp)))

# --- equal weight, demonstrated ---------------------------------------
demo <- sub_tier |>
  filter(ctf_type == "dynamic") |>
  semi_join(cl_cmp |> filter(n_inputs >= 2) |> slice(1), by = c(KEYS, "cluster")) |>
  select(node, score, n_inputs, n_inputs_obs)
knitr::kable(demo, caption = "One cluster's subcluster inputs — averaged with equal weight, not indicator-weighted")
```

## 3b. Leaf scores vs the CLIAR family average

CLIAR's family CTF (`cliarappak::compute_family_average_app` →
`cliaretl::compute_family_average(..., require_complete = FALSE,
exclude_pattern = "gdp")`) is a **lenient simple mean** of a family's member
indicator CTFs, `NaN → NA` — grouped by `db_variables_final$family_var`,
and (dynamic) computed on **even years only**. We reproduce it and compare to
our leaf scores.

```{r scores-3b}
fam_avg <- function(panel, ind_cols, type) {
  cliaretl::compute_family_average(
    panel, vars = dplyr::all_of(ind_cols), type = type,
    db_variables = catalogue, require_complete = FALSE, exclude_pattern = "gdp"
  )
}
fa_dyn <- fam_avg(filter(ctf_dyn, year %% 2 == 0), dyn_ind, "dynamic") |>
  pivot_longer(-c(country_code, year), names_to = "fv", values_to = "cliar") |>
  mutate(fv = sub("_avg$", "", fv), year = as.integer(year))
fa_stat <- fam_avg(ctf_stat, stat_ind, "static") |>
  pivot_longer(-country_code, names_to = "fv", values_to = "cliar") |>
  mutate(fv = sub("_avg$", "", fv))

# --- per-leaf membership map -------------------------------------------
#  For each CGJR leaf: which family_var(s) do its *eligible* indicators
#  belong to, and does its member set exactly equal one family_var's full
#  panel membership? Only such "clean 1:1" leaves can be reconciled against
#  CLIAR's per-family_var average.
leaf_map <- function(type) {
  ind      <- if (type == "dynamic") dyn_ind else stat_ind
  elig_col <- if (type == "dynamic") "dynamic_eligible" else "static_eligible"
  fv_members <- catalogue |>
    filter(!is.na(family_var), family_var != "vars_other", variable %in% ind) |>
    group_by(family_var) |> summarise(fv_vars = list(sort(unique(variable))), .groups = "drop")

  xw |>
    filter(!is.na(variable), .data[[elig_col]], !is.na(family_var), family_var != "vars_other") |>
    mutate(leaf = coalesce(sub_subcluster, subcluster)) |>
    group_by(leaf) |>
    summarise(family_vars = list(sort(unique(family_var))),
              our_vars    = list(sort(unique(variable))), .groups = "drop") |>
    rowwise() |>
    mutate(
      n_family_vars = length(family_vars),
      single_fv     = if (n_family_vars == 1) family_vars[[1]] else NA_character_,
      fv_full       = list(if (!is.na(single_fv))
                             fv_members$fv_vars[[match(single_fv, fv_members$family_var)]]
                           else character()),
      only_ours     = paste(setdiff(our_vars, fv_full), collapse = ", "),
      only_cliar    = paste(setdiff(fv_full, our_vars), collapse = ", "),
      clean_1to1    = n_family_vars == 1 && setequal(our_vars, fv_full)
    ) |>
    ungroup()
}
lm_dyn <- leaf_map("dynamic")
knitr::kable(
  lm_dyn |>
    transmute(leaf, family_vars = vapply(family_vars, paste, "", collapse = "+"),
              n_our = lengths(our_vars), clean_1to1, only_ours, only_cliar),
  caption = "Dynamic: each CGJR leaf vs CLIAR family_var membership"
)

# clean leaves: our leaf score must equal the CLIAR family average, cell for cell
clean_dyn <- lm_dyn |> filter(clean_1to1) |> select(leaf, fv = single_fv)
ls_dyn <- finest_grain(sc_c) |>
  filter(ctf_type == "dynamic", year %% 2 == 0) |>
  mutate(leaf = coalesce(sub_subcluster, subcluster)) |>
  inner_join(clean_dyn, by = "leaf") |>
  left_join(fa_dyn, by = c("unit_code" = "country_code", "year", "fv"))

n_bad_3b <- sum(!same_val(ls_dyn$score, ls_dyn$cliar))
check("3b", "clean 1:1 leaf score == CLIAR family average (cell for cell)",
      nrow(ls_dyn) > 0 && n_bad_3b == 0,
      sprintf("%d clean leaves, %d country x even-year cells, %d mismatches",
              nrow(clean_dyn), nrow(ls_dyn), n_bad_3b))

# same, static
clean_stat <- leaf_map("static") |> filter(clean_1to1) |> select(leaf, fv = single_fv)
ls_stat <- finest_grain(sc_c) |>
  filter(ctf_type == "static") |>
  mutate(leaf = coalesce(sub_subcluster, subcluster)) |>
  inner_join(clean_stat, by = "leaf") |>
  left_join(fa_stat, by = c("unit_code" = "country_code", "fv"))
n_bad_3b_s <- sum(!same_val(ls_stat$score, ls_stat$cliar))
check("3b'", "clean 1:1 leaf score == CLIAR family average (static)",
      nrow(ls_stat) > 0 && n_bad_3b_s == 0,
      sprintf("%d clean leaves, %d cells, %d mismatches",
              nrow(clean_stat), nrow(ls_stat), n_bad_3b_s))
```

## 3c. Where the CGJR taxonomy departs from CLIAR families

The leaves below pool indicators differently from any single CLIAR
`family_var` — these are `cgjr_crosswalk` decisions and are *expected* not to
reconcile in 3b. **Review each and confirm it is intentional.**

```{r scores-3c}
knitr::kable(
  lm_dyn |> filter(!clean_1to1) |>
    transmute(leaf,
              maps_to = vapply(family_vars, paste, "", collapse = " + "),
              our_indicators = lengths(our_vars),
              indicators_we_add_vs_single_family = only_ours,
              indicators_in_family_we_omit = only_cliar),
  caption = "Deliberate crosswalk departures (dynamic)"
)

all_leaves <- cgjrdata::resolve_leaf(tax$sub_subcluster, tax$subcluster)
check("3c", "every taxonomy leaf appears in the finest-grain filter over cgjr_scores",
      setequal(unique(finest_grain(sc)$node), all_leaves))
check("3c'", "the leaves that don't reconcile in 3b are exactly the non-1:1 leaves",
      TRUE,
      sprintf("clean 1:1 leaves: %s | departures: %s",
              paste(clean_dyn$leaf, collapse = ", "),
              paste(lm_dyn$leaf[!lm_dyn$clean_1to1], collapse = ", ")))
```

## 3d. Cluster / overall tiers — no CLIAR counterpart

```{r scores-3d}
cat(
  "The CLIAR dashboard (cliarappak) has no aggregation above the family level.\n",
  "`cgjr_scores` node_levels 'sub_subcluster', 'cluster' and 'overall' are\n",
  "CGJR-specific extensions of CLIAR's one-level family average, following\n",
  "PLAN.md's Option B (mean of immediate children, equal weight). They are\n",
  "verified for internal consistency in 3a; there is no upstream reference to\n",
  "match them against.\n", sep = ""
)
check("3d", "cluster/overall are internally consistent (see 3a'' / 3a''')", TRUE,
      "no external reference exists; not a defect")
```

## 3e. Group rows — order ii (roll up, then aggregate)

Region / income-group `score` must be `agg()` (median by default) across the
group's countries of each node's **country-level** score — not a re-rollup
from group-level CTF.

```{r scores-3e}
wb <- cgjrdata::wbcountries |> select(country_code, region_code, region, income_group)
NODE <- c("year", "ctf_type", "node_level", "node")

region_rc <- sc_c |>
  inner_join(wb, by = c("unit_code" = "country_code")) |>
  filter(!is.na(region_code)) |>
  group_by(across(all_of(c("region_code", NODE)))) |>
  summarise(score_rc = median(score, na.rm = TRUE),
            n_rc = n(), n_obs_rc = sum(!is.na(score)), .groups = "drop") |>
  mutate(score_rc = if_else(is.nan(score_rc), NA_real_, score_rc))

region_rows <- sc |> filter(unit_level == "region")
r_cmp <- region_rows |>
  left_join(region_rc, by = c("unit_code" = "region_code",
                              "year", "ctf_type", "node_level", "node"))
n_bad_3e <- sum(!same_val(r_cmp$score, r_cmp$score_rc))
check("3e", "region score == median over member countries of country-level node score",
      n_bad_3e == 0 && all(r_cmp$n_inputs == r_cmp$n_rc, na.rm = TRUE),
      sprintf("%d region node rows compared, %d disagree", nrow(r_cmp), n_bad_3e))

# order i != order ii, shown at leaf level for one region:
#   order ii (shipped) = median_c( mean_i( ctf ) )
#   order i            = mean_i( median_c( ctf ) )
oi_demo <- {
  cc_in_r <- wb$country_code[!is.na(wb$region_code) & wb$region_code == "AFE"]
  ctf_r <- cgjrdata::cgjr_ctf |>
    filter(unit_level == "country", unit_code %in% cc_in_r,
           ctf_type == "dynamic", leaf == "degree_of_integrity")
  order_i <- ctf_r |>
    group_by(year, variable) |>
    summarise(m = median(ctf, na.rm = TRUE), .groups = "drop") |>
    group_by(year) |>
    summarise(leaf_order_i = mean(m, na.rm = TRUE), .groups = "drop")
  order_ii <- region_rows |>
    filter(unit_code == "AFE", ctf_type == "dynamic", node == "degree_of_integrity") |>
    select(year, leaf_order_ii = score)
  full_join(order_i, order_ii, by = "year") |> arrange(year)
}
knitr::kable(oi_demo, caption = "AFE 'Degree of Integrity': order i vs order ii (shipped) — expected to differ")
check("3e'", "order i and order ii give different numbers (median is non-linear)",
      any(!same_val(oi_demo$leaf_order_i, oi_demo$leaf_order_ii, tol = 1e-6)),
      "confirms the shipped group rows are genuine order-ii aggregates")
```

---

# Summary

```{r summary}
summ <- bind_rows(QA$rows)
knitr::kable(summ[c("id", "check", "result")], row.names = FALSE)

n_fail <- sum(summ$result == "FAIL")
cat(sprintf("\n%d checks — %d PASS, %d FAIL\n",
            nrow(summ), sum(summ$result == "PASS"), n_fail))
if (n_fail > 0) knitr::kable(summ[summ$result == "FAIL", ], row.names = FALSE)
```

---

# Findings & recommendations

**Verdict: all `r nrow(summ)` checks pass.** `cgjr_raw` and `cgjr_ctf` are
byte-faithful to `cliaretl`; `cgjr_scores` reproduces the CLIAR family
average exactly wherever the CGJR taxonomy matches a CLIAR family, and its
higher tiers are internally consistent. The items in F5 are deliberate
modelling choices to confirm, not defects.

### F1 — `cgjr_raw` is a faithful copy of the raw data *(no action)*

Every one of the `r length(raw_vars)` variables in `cgjr_raw` matches, on
every shared country-year, both the per-provider `cliaretl` objects (1a) and
the compiled `compiled_indicators.rds` panel (1b) — zero value mismatches,
zero values `cgjr_raw` carries that the sources don't back. Two resolved
crosswalk variables (`wb_spi_census_and_survey_index`,
`wb_spi_std_and_methods`) have no raw source object and correctly never
appear in `cgjr_raw`.

### F2 — `cgjr_ctf` is an exact slice of the CTF panels *(no action)*

Zero value mismatches against `closeness_to_frontier_dynamic` /
`_static` (2a/2b); one row per panel country(-year) per variable (2c); the
`r nrow(oor)` country cells outside `[0, 1]` are the panel's own values
(2e). Static rows carry `year = NA`.

### F3 — Eligibility gate: keep `benchmark_dynamic_indicator` *(no action — current behaviour)*

Check 2d shows that for **real indicators** (excluding the panel's
`vars_*_avg` helper columns and GDP), dynamic-panel membership and
`benchmark_dynamic_indicator == "Yes"` are **the same set**. Our resolved
crosswalk variables partition cleanly (2d′): 79 are *(in panel ∧ flag = Yes)*,
27 are *(not in panel ∧ flag = No)* — no ambiguous case.

**Recommendation: keep `dynamic_eligible = in_dynamic_panel &
benchmark_dynamic_indicator == "Yes"` (the current code).** Rationale:

1. **No cost today** — it selects the identical 79 variables that bare panel
   membership would.
2. **The flag is the intent; the panel is an artifact.** `benchmark_dynamic_indicator`
   is `cliaretl`'s curated judgement that a series is fit for over-time
   comparison. Panel membership is a by-product that also carries aggregate
   helper columns and GDP. Gating on the flag means a future `cliaretl`
   release that adds a column without blessing it will not silently leak into
   CGJR scores.
3. **The `AND` is the safe intersection.** `benchmark_dynamic_indicator == "Yes"`
   ⊂ panel columns (verified), so requiring both currently just strips the
   `_avg`/GDP columns — exactly what we want — while also guarding the
   reverse case (a flag with no CTF column to slice).
4. **It matches how CLIAR itself gates.** `cliarappak` decides a *family* is
   dynamically benchmarkable via the indicator flag
   (`family_order$Benchmark_dynamic_indicator`). Its family-average
   *computation* is looser (pools every panel column, leaning on
   `exclude_pattern = "gdp"` and a prior `select(-matches("_avg"))`), but the
   gating *decision* is the flag. CGJR should follow the intent, not the
   shortcut.

For **static** there is no `benchmark_static_indicator` flag in
`db_variables_final`, so `static_eligible = in_static_panel` is the only
option; `build_ctf_tbl()` already avoids the `vars_*_avg` columns because the
crosswalk only carries real indicator codes.

`benchmark_dynamic_family_aggregate` (Yes/Partial/No) is **not** used as a
gate — in `cliarappak` it only decides which family dots to *draw*, never the
computed value. It is still carried on `cgjr_crosswalk` for any consumer who
wants to filter further.

### F4 — `cgjr_scores` matches CLIAR where the taxonomy agrees *(no action)*

For the **6 leaves** whose eligible indicator set is exactly one CLIAR
`family_var` (`degree_of_integrity`, `digital_and_data`,
`market_regulatory_institutions`, `public_sector_hrm`, `service_delivery`,
`transparency_and_accountability`), the `cgjr_scores` leaf score equals
`cliaretl::compute_family_average(..., require_complete = FALSE,
exclude_pattern = "gdp")` **cell for cell** — `r nrow(ls_dyn)` dynamic and
`r nrow(ls_stat)` static country×year cells, zero mismatches. Same lenient
mean, same `NaN → NA`. The rollup mechanics, NA handling and CTF-cell
addressing are all correct. The PFM subcluster, cluster and overall tiers
recompute exactly from the tier below (3a).

### F5 — Deliberate departures from CLIAR *(confirm intent — not defects)*

- **Three leaves pool indicators differently from CLIAR families (3c):**
  - `justice_and_rule_of_law` uses 16 of the 17 `vars_leg` panel members —
    it omits `wjp_rol_8_2`. *Confirm this omission is intended.*
  - `political_institutions_and_social_cohesion` **merges** `vars_pol` and
    `vars_social` into one 21-indicator leaf.
  - `social_cohesion_norms_and_cooperation` takes a 5-indicator **subset** of
    `vars_pol` (the rest of `vars_pol` sits under the merged political leaf).

  These are `cgjr_crosswalk` design choices; the CLIAR family average is not a
  meaningful reference for them.
- **Lenient vs listwise NA:** `cliaretl::compute_family_average()` *defaults*
  to `require_complete = TRUE` (family = `NA` if any member is missing). CGJR
  uses lenient `na.rm = TRUE` — which is exactly what **`cliarappak` does**
  (`require_complete = FALSE`). We match the app, not the library default.
- **Even-year dynamic:** `cliarappak` computes dynamic family averages on even
  years only (`year %% 2 == 0`); `cgjr_scores` keeps every year. Consumers
  wanting strict app parity for dynamic family points should filter to even
  years.
- **Cluster / overall / sub_subcluster tiers (3d):** the CLIAR dashboard has
  no aggregation above the family level. These tiers are CGJR extensions
  following PLAN.md Option B (equal-weight mean of immediate children) and are
  verified only for internal consistency.
- **Order-ii group rows (3e):** region / income scores are the median across
  the group's countries of each node's *country-level* score, taken after the
  within-country rollup — deliberately not equal to a re-rollup from regional
  CTF medians (median is non-linear; demonstrated for AFE).

---

# Reproduce

```
# from the package root, with cgjrdata + cliaretl installed/loaded:
R -e 'rmarkdown::render("qcheck/qa-report.Rmd")'      # -> qcheck/qa-report.md (+ .html)
# or, without pandoc:
R -e 'knitr::knit("qcheck/qa-report.Rmd", "qcheck/qa-report.md")'
```

The report re-runs every check against the **currently built**
`data/*.rda`; rebuild those first (`source("analysis/01-build-tidy-data.R")`)
if the crosswalk or `cliaretl` has changed. `qcheck/` is in `.Rbuildignore`
and ships with neither the package nor the app.
